Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-08-04 doc/GMP2015/revision_3/s02_preliminaries.tex
Made a talk, consolidated doc/img, began new revision, updated doc todo and .gitignore
Updated paper to say algo is O(V^4). However, I also provide a sketch of what I believe to be a working O(V^3). Updated todo's.

% \ClearMyMinHeight
% \SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}

% \begin{figure*}\centering%
%   %
%   \begin{subfigure}{0.4\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
%     \caption{}\label{fig:demo_graph:graph}
%   \end{subfigure}%
%   %
%   \hfill
%   \begin{subfigure}{0.3\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
%     \caption{}\label{fig:demo_graph:comdrp}
%   \end{subfigure}%
%   %
%   \hfill
%   \begin{subfigure}{0.3\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
%     \caption{}\label{fig:demo_graph:candrp}
%   \end{subfigure}%
%   %
%   \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$. The children of triangles are omitted.}
%   \label{fig:demo_graph}
% \end{figure*}%

% \begin{figure*}\centering%
%   %
%   \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
%   \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality. Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
%   \label{fig:demo_graph:clustmindrp}
% \end{figure*}%

\ClearMyMinHeight
\SetMyMinHeight{.4}{img/svg/3xc2c3}
\SetMyMinHeight{.3}{img/svg/3xc2c3_comdrp}
\SetMyMinHeight{.3}{img/svg/3xc2c3_candrp}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.4\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3}
    \caption{}\label{fig:demo_graph:graph}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_comdrp}
    \caption{}\label{fig:demo_graph:comdrp}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_candrp}
    \caption{}\label{fig:demo_graph:candrp}
  \end{subfigure}%
  %
  \caption{(\ref{fig:demo_graph:graph}) A graph, $G_{demo}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{demo}$, i.e.\ $ComDRP(G_{demo})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{demo}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{demo})$. The children of triangles are omitted.}
  \label{fig:demo_graph}
\end{figure*}%

% \ClearMyMinHeight
% \SetMyMinHeight{.4}{img/svg/triangle}
% \SetMyMinHeight{.3}{img/svg/triangle_candrp}
% \SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}

%\begin{figure*}\centering%
%
%\begin{subfigure}{0.4\linewidth}\centering
%\includegraphics[height=\myMinHeight]{img/svg/triangle}
%    \caption{}\label{fig:demo_graph_tri:graph}
% \end{subfigure}%
%
%  \hfill
% \begin{subfigure}{0.3\linewidth}\centering
%   \includegraphics[height=\myMinHeight]{img/svg/triangle_candrp}
%    \caption{}\label{fig:demo_graph_tri:candrp}
%  \end{subfigure}%
%
%  \hfill
%  \begin{subfigure}{0.3\linewidth}\centering
%   \includegraphics[height=\myMinHeight]{img/svg/triangle_clustmindrp}
%   \caption{}\label{fig:demo_graph_tri:clustmindrp}
%  \end{subfigure}%
%
%  \caption{(\ref{fig:demo_graph_tri:graph}) A simple graph, $G_{tri}$, of 3 triangles intersecting trivially. (\ref{fig:demo_graph_tri:candrp}) The canonical DR-plan of $G_{tri}$. (\ref{fig:demo_graph_tri:clustmindrp}) A DR-plan of graph $G_{tri}$ which gives cluster minimality  (See Section \ref{sec:prev}). Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the canonical/optimal DR-plan has fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
% \label{fig:demo_graph:clustmindrp}
%  \label{fig:demo_graph_tri}
%\end{figure*}%

\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, is defined as a forest that has the following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item The children $C_1,\ldots,C_N$ of a node $C$ satisfy $\bigcup_{i=1}^N{C_i}=C$.
        \item A leaf node is a single edge. A \dfn{trivial} graph is empty or a single vertex. Note that a trivial graph is not isostatic.
        \item A root node is a vertex-maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$. This makes Property 2 implicit.
  %and in fact, $C$ contains all the edges in the union of its children as well.
  We denote a complete DR-plan of $G$ as $\comdrp{G}$.

  A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree. The maximum  fan-in is called the \dfn{size} of the DR-plan. We denote an optimal DR-plan of $G$ as $\optdrp{G}$.
%
%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
%an isostatic vertex-maximal proper subgraph.
    %Note that nodes will be referred to interchangeably as ``the node that
    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
\end{definition}
%
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
For a given graph, there could be exponentially many DR-plans---and even optimal DR-plans---in the size of the graph. A complete DR-plan is unique but may not be (and is usually not) optimal. DR-plans of self-similar graphs are self-similar.
\end{remark}

@@

% \begin{definition}
%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
% \end{definition}

@@

%    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.

@@

%%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.

>>>>

% \ClearMyMinHeight
% \SetMyMinHeight{.4}{../../img/epsfromtikz/demo_graph}
% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_comdrp}
% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_candrp}

% \begin{figure*}\centering%
%   %
%   \begin{subfigure}{0.4\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph}
%     \caption{}\label{fig:demo_graph:graph}
%   \end{subfigure}%
%   %
%   \hfill
%   \begin{subfigure}{0.3\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph_comdrp}
%     \caption{}\label{fig:demo_graph:comdrp}
%   \end{subfigure}%
%   %
%   \hfill
%   \begin{subfigure}{0.3\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/demo_graph_candrp}
%     \caption{}\label{fig:demo_graph:candrp}
%   \end{subfigure}%
%   %
%   \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$. The children of triangles are omitted.}
%   \label{fig:demo_graph}
% \end{figure*}%

% \begin{figure*}\centering%
%   %
%   \includegraphics[width=0.4\linewidth]{../../img/epsfromtikz/demo_graph_clustmindrp}
%   \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality. Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
%   \label{fig:demo_graph:clustmindrp}
% \end{figure*}%

\ClearMyMinHeight
\SetMyMinHeight{.4}{../../img/svg/3xc2c3}
\SetMyMinHeight{.3}{../../img/svg/3xc2c3_comdrp}
\SetMyMinHeight{.3}{../../img/svg/3xc2c3_candrp}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.4\linewidth}\centering
    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3}
    \caption{}\label{fig:demo_graph:graph}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_comdrp}
    \caption{}\label{fig:demo_graph:comdrp}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{../../img/svg/3xc2c3_candrp}
    \caption{}\label{fig:demo_graph:candrp}
  \end{subfigure}%
  %
  \caption{(\ref{fig:demo_graph:graph}) A graph, $G_{demo}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{demo}$, i.e.\ $ComDRP(G_{demo})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{demo}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{demo})$. The children of triangles are omitted.}
  \label{fig:demo_graph}
\end{figure*}%

% \ClearMyMinHeight
% \SetMyMinHeight{.4}{../../img/svg/triangle}
% \SetMyMinHeight{.3}{../../img/svg/triangle_candrp}
% \SetMyMinHeight{.3}{../../img/svg/triangle_clustmindrp}

%\begin{figure*}\centering%
%
%\begin{subfigure}{0.4\linewidth}\centering
%\includegraphics[height=\myMinHeight]{../../img/svg/triangle}
%    \caption{}\label{fig:demo_graph_tri:graph}
% \end{subfigure}%
%
%  \hfill
% \begin{subfigure}{0.3\linewidth}\centering
%   \includegraphics[height=\myMinHeight]{../../img/svg/triangle_candrp}
%    \caption{}\label{fig:demo_graph_tri:candrp}
%  \end{subfigure}%
%
%  \hfill
%  \begin{subfigure}{0.3\linewidth}\centering
%   \includegraphics[height=\myMinHeight]{../../img/svg/triangle_clustmindrp}
%   \caption{}\label{fig:demo_graph_tri:clustmindrp}
%  \end{subfigure}%
%
%  \caption{(\ref{fig:demo_graph_tri:graph}) A simple graph, $G_{tri}$, of 3 triangles intersecting trivially. (\ref{fig:demo_graph_tri:candrp}) The canonical DR-plan of $G_{tri}$. (\ref{fig:demo_graph_tri:clustmindrp}) A DR-plan of graph $G_{tri}$ which gives cluster minimality  (See Section \ref{sec:prev}). Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the canonical/optimal DR-plan has fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
% \label{fig:demo_graph:clustmindrp}
%  \label{fig:demo_graph_tri}
%\end{figure*}%

% \begin{definition}\label{def:drp}
%     The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, is defined as a forest that has the following properties:
%     \begin{enumerate}
%         \item Each node represents/contains/is a rigid subgraph of $G$.
%         % \item For a node, $C$, that is  a non-trivial graph, its $N$
%         % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
%         % subgraphs of $C$.
%         \item The children $C_1,\ldots,C_N$ of a node $C$ satisfy $\bigcup_{i=1}^N{C_i}=C$.
%         \item A leaf node is a single edge. A \dfn{trivial} graph is empty or a single vertex. Note that a trivial graph is not isostatic.
%         \item A root node is a vertex-maximal rigid subgraph of $G$.
%     \end{enumerate}
% %
%     % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
% %
%     A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$. This makes Property 2 implicit.
%   %and in fact, $C$ contains all the edges in the union of its children as well.
%   We denote a complete DR-plan of $G$ as $\comdrp{G}$.

%   A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the forest. The maximum  fan-in is called the \dfn{size} of the DR-plan. We denote an optimal DR-plan of $G$ as $\optdrp{G}$.
% %
% %In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
% %$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
% %an isostatic vertex-maximal proper subgraph.
%     %Note that nodes will be referred to interchangeably as ``the node that
%     %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
% \end{definition}

% We begin with the definition of a general DR-plan and also introduce two specific kinds of DR-plans.

\begin{definition}\label{def:drp}
    A \dfn{decomposition-recombination \mbox{(DR-)} plan}~\cite{hoffman2001decompositionI} of graph $G$ is defined as a forest that has the following properties:
    \begin{enumerate}
        \item Each node represents a rigid subgraph of $G$.
        \item A root node is a vertex-maximal rigid subgraph of $G$.
        \item A node is the subgraph of $G$ induced by the union of its children.
        \item A leaf node is a single edge.
    \end{enumerate}
\end{definition}

\begin{definition}
    A DR-plan is \dfn{complete} if it satisfies an additional property: for a non-leaf node $C$, its children are all of the rigid vertex-maximal proper subgraphs of $C$ (which makes Property 3 of a DR-plan implicit.)
\end{definition}

\begin{definition}
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the forest. The maximum fan-in is called the \dfn{size} of the DR-plan.
\end{definition}

We denote these different kinds of DR-plans as $\drp{G}$, $\comdrp{G}$, and $\optdrp{G}$, respectively. Note that nodes will be referred to interchangeably as ``the node that represents or contains the (sub)graph $C$'' and as simply ``$C$''.

%
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
More than one node (leaf) in a DR-plan forest may represent the same subgraph (vertex) of $G$.
For a given graph, there could be exponentially many DR-plans---and even optimal DR-plans---in the size of the graph. A complete DR-plan is unique but may not be (and is usually not) optimal. DR-plans of self-similar graphs are self-similar.
\end{remark}

@@

% \begin{definition}
%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the forest. It is not necessarily unique.
% \end{definition}

@@

%    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the forest. It is not necessarily unique.

@@

%%     The \textbf{optimal DR-plan} of graph $G$, $OptimalDRP(G)$, is the DR-plan that minimizes the maximum fan-in over all nodes in the forest. It is not necessarily unique.
